Inducing Semantic Frames on a Very Large Corpus of Syntactic-Ngrams

### 2014-12-21 Paper/paper.tex
Model 1.

\begin{figure}[h]

@@

    \label{gen0}
    \caption{Model 0 generative story}

@@

The generative story (fig. \ref{gen0}) gives us the following joint probability:
\begin{align*}
&P(\mathbf{f},\mathbf{w}|\phi,\theta) 
  = P(\mathbf{f}|\theta)P(\mathbf{w}|\phi)\\
  =& \prod_{i=1}^{N}\big[\theta(f_i) \prod_a^{\{v,s,o\}}\phi_{f_i}^a(w_i^a)\big]
\end{align*}

@@

As we know from topic modeling, it is the nature of documents that they contain 
sentences with related semantic content. Therefore, it is reasonable to assume 
that semantic frames are not uniformly distributed within documents. Taking 
advantage of this assumption, we can break drop model 0's assumption of independence 
between tuples.

@@

    \caption{Model 1 generative story}

@@

The independence assumptions int the generative story for model 1 give us the 
following joint distribution:

@@

\begin{align}
&p(z_{ij} = f| \mathbf{f}_{-ij},\alpha,\beta,\mathbf{w})\\
&\propto \prod_a^{\{v,s,o\}}\frac{\beta^a + \tilde{c}(f,w_{ij})}{V^a + \tilde{c}(f)}
        \cdot \frac{\alpha + \tilde{c}(i,f)}{f\alpha+\tilde{c}(i)}
\end{align}

>>>>

@@

    \caption{Model 0 generative story}
    \label{gen0}

@@

The generative story (figure \ref{gen0}) gives us the following joint probability:
\begin{align*}
&P(\mathbf{f},\mathbf{w}|\phi,\theta) 
  = P(\mathbf{f}|\theta)P(\mathbf{w}|\phi)\\
  =& \prod_{i=1}^{N}\big[\theta(f_i) \prod_a^{\{v,s,o\}}\phi_{f_i}^a(w_i^a)\big]
\end{align*}

@@

It is in the nature of documents that they have some degree of internal semantic
coherence. Put more plainly, a document is usually \emph{about} something. We
know from topic modeling that LDA with Gibbs sampling can be very successful at
classifying documents according to some semantic features \citep{griffiths2004}.
Model 1 leverages this fact by making the assmumption that a given document is 
likely to have a sparse distribution of frames. As in LDA, the document-level
sparsity is enforced by a dirichlet prior on frame distributions (see figure
\ref{gen1} for details). This assumption breaks the independence between tuples 
that is maintained by Model 0 is dropped in Model 1.

@@

    \caption{Model 1 generative story}
    \label{gen1}

@@

The independence assumptions in the generative story for model 1 give us the 
following joint distribution:

@@

These integrals are both dirichlet multinomials:

\[
\frac{DCM_i^{frames}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{frames}(\mathbf{f}_{-ij})}
\]

\[
\prod_a^{\{v,s,o\}}\frac{DCM_i^{vocab^a}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{vocab^a}(\mathbf{f}_{-ij})}
\]

We prefer not to consider the subject and object vocabularies independently 
(i.e., $V^s = V^o$). This is more in line with the theory of semantic frames 
since in principle, a noun playing a particlar semantic role may appear in either 
syntactic position.

Dropping terms not dependent on counts, we can calculate the probability 
required for Gibbs sampling as follows:

\begin{align}
& P(f_{ij} = f|\mathbf{f}_{-ij},\mathbf{w}, \alpha,\beta)\nonumber\\
&=\frac{\alpha + \tilde c(d_i,f)}{F\,\alpha + \tilde c(d_i)}
\times \prod_a^{\{v,s,o\}}\frac{\beta+\tilde c(f,w_{ij}^a)}{V^a\beta+\tilde c(f)}
\end{align}

Where $\tilde c (d_i, f)$ is the number of VSO's in document $d_i$ assigned to frame $f$,
$\tilde c(d_i)$ is the number of VSO's in document $i$,
$\tilde c(f,w_{ij}^a)$ is the number of VSOs containing $w_{ij}^a$ as argument $a$ that are
assigned to frame $f$, and
$\tilde c(f)$ is the number of VSOs assigned to frame $f$.
Note that $\tilde c$ indicates a count that does not consider the contribution of 
VSO $w_{i,j}$.

We will see that the data used in this probject doesn't contain any document-level
information. As a result, we adapt model 1 by grouping VSOs with the same verb into ``documents''.
This move leverages the dirichlet prior $\alpha$ to enforce sparcity of frames 
accross verbs, but of course cannot capture the information given by the kind of
document-level semantic sparsity seen in LDA.\footnote{When we use verbs as
documents (i.e, $D = V^v$), since $w_{ij}^v$ is the same for all verbs in document $d_i$ 
(and $d_i$ is the only document with that verb), it follows that the number of VSO's in document 
$d_i$ belonging to frame $f$ is the same as the number of VSO's with $w_{ij}^v$ as their verb assigned to 
frame $f$; that is, $\tilde c(d_i, f) = \tilde c(f, w_{ij}^v)$.}
